---
title: "Critical evaluation of the reliability of DNA methylation probes on the Illumina MethylationEPIC BeadChip microarrays for dementia research"
subtitle: "ICC vs ADNI study results"
author: "Wei Zhang, Lily Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: yeti
    toc: true
    number_sections: true
    df_print: paged
    code_download: false
    toc_float:
      collapsed: yes
    toc_depth: 3
    font_size: 10
editor_options:
  chunk_output_type: inline 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(gmodels)

dir.base <- "../../"
dir.reliable.score <- file.path(dir.base, "analysis-results/reliable_score_different_plate")
dir.plot <- file.path(dir.base, "analysis-results/plots")

dir.mike <- file.path(dir.base, "Mike")
dir.extra <- file.path(dir.base, "external_files")
```

# Demonstrate reliable CpGs are more consistent across studies

```{r message = F}
reliable_df <- read_csv(file.path(dir.reliable.score, "ADNI_icc_beta_and_M_results_ind_samples_annotated_different_plates.csv"))
```

## Sex-combined study

```{r eval = F}
### Load results
adni.results <- read_csv(file.path(dir.extra, "meta_analysis_glm_fixed_effect_ADNI_and_AIBL_AD_vs_CN_single_cpg_annotated.csv"),
                         show_col_types = FALSE)

reliable_df <- inner_join(
  reliable_df,
  adni.results %>% dplyr::select(cpg, estimate.bacon, se.bacon, pVal.final.bacon, direction.bacon,
                                 ADNI_AD_vs_CN_Estimate.bacon, ADNI_AD_vs_CN_pValue.bacon,
                                 AIBL_AD_vs_CN_Estimate.bacon, AIBL_AD_vs_CN_pValue.bacon)
)
reliable_df$effect_size_diff <- 
  abs(reliable_df$ADNI_AD_vs_CN_Estimate.bacon - reliable_df$AIBL_AD_vs_CN_Estimate.bacon)
write_csv(
  reliable_df,
  file.path(dir.reliable.score, "ADNI_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```


```{r message = F}
reliable_df <- read_csv(
  file.path(dir.reliable.score, "ADNI_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_df,
  y = "effect_size_diff",
  x = "ICC_Beta",
  ylab = "Effect size absolute differences between ADNI and AIBL",
  xlab = "ICC",
  size = .1,
  cor.coef = T,
  cor.method = "spearman"
)
```


```{r  echo = F, fig.width=9, fig.height=8}
ggboxplot(
  data = reliable_df, 
  x = "group",
  y = "effect_size_diff",
  ylab = "Effect size absolute differences between ADNI and AIBL",
  xlab = "ICC and HoLA classification",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA,
  ylim = c(0, 45)
) 
```

```{r echo = F, fig.width=9, fig.height=8}
ggboxplot(
  data = reliable_df, 
  x = "classification_Beta",
  y = "effect_size_diff",
  ylab = "Effect size absolute differences between ADNI and AIBL",
  xlab = "ICC group",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA,
  ylim = c(0, 45)
) 
```

## Sex-specific study

```{r message = F}
reliable_df <- read_csv(file.path(dir.reliable.score, "ADNI_icc_beta_and_M_results_ind_samples_annotated_different_plates.csv"))
```

### Female

```{r eval = F}
female.results <- read_csv(
  file.path(dir.reliable.score, "../effect_size/FEMALE_meta_analysis_glm_fixed_effect_ADNI_and_AIBL_AD_vs_CN_single_cpg_annotated.csv")
)
reliable_df <- inner_join(
  reliable_df,
  female.results %>% dplyr::select(cpg,  ADNI_FEMALE_Estimate.bacon, ADNI_FEMALE_pValue.bacon, AIBL_FEMALE_Estimate.bacon, AIBL_FEMALE_pValue.bacon)
)
reliable_df$effect_size_diff <- abs(
  reliable_df$ADNI_FEMALE_Estimate.bacon - reliable_df$AIBL_FEMALE_Estimate.bacon)
write_csv(
  reliable_df,
  file.path(dir.reliable.score, "FEMALE_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```



```{r message = F}
reliable_df <- read_csv(
  file.path(dir.reliable.score, "FEMALE_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_df,
  y = "effect_size_diff",
  x = "ICC_Beta",
  ylab = "Female effect size abs differences between ADNI and AIBL",
  xlab = "ICC",
  size = .1,
  cor.coef = T,
  cor.method = "spearman"
)
```


```{r echo = F, fig.width=9, fig.height=8}
ggboxplot(
  data = reliable_df, 
  x = "classification_Beta",
  y = "effect_size_diff",
  ylab = "Female effect size absolute differences between ADNI and AIBL",
  xlab = "ICC group",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA,
  ylim = c(0, 1)
) 
```


### Male

```{r message = F}
reliable_df <- read_csv(file.path(dir.reliable.score, "ADNI_ICC_annotation_with_HoLA_grouped.csv"))
```


```{r eval = F}
male.results <- read_csv(
  file.path(dir.reliable.score, "../effect_size/MALE_meta_analysis_glm_fixed_effect_ADNI_and_AIBL_AD_vs_CN_single_cpg_annotated.csv")
)
reliable_df <- inner_join(
  reliable_df,
  male.results %>% dplyr::select(cpg, 
                                 ADNI_MALE_Estimate.bacon, ADNI_MALE_pValue.bacon,
                                 AIBL_MALE_Estimate.bacon, AIBL_MALE_pValue.bacon)
)
reliable_df$effect_size_diff <- abs(
  reliable_df$ADNI_MALE_Estimate.bacon - reliable_df$AIBL_MALE_Estimate.bacon)
write_csv(
  reliable_df,
  file.path(dir.reliable.score, "MALE_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```


```{r message = F}
reliable_df <- read_csv(
  file.path(dir.reliable.score, "MALE_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL.csv")
)
```

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_df,
  y = "effect_size_diff",
  x = "ICC_Beta",
  ylab = "Male effect size abs differences between ADNI and AIBL",
  xlab = "ICC",
  size = .1,
  cor.coef = T,
  cor.method = "spearman"
)
```

```{r}
ggboxplot(
  data = reliable_df, 
  x = "classification_Beta",
  y = "effect_size_diff",
  ylab = "Male effect size absolute differences between ADNI and AIBL",
  xlab = "ICC group",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA,
  ylim = c(0, 1)
) 
```


# Demonstrate reliable CpGs have methylation levels that are more consistent across brain and blood 

```{r message = F}
reliable_df <- read_csv(file.path(dir.reliable.score, "ADNI_icc_beta_and_M_results_ind_samples_annotated_different_plates.csv"))
```

```{r message = F}
london.corr <- read_csv(
  file.path(dir.extra, "London_blood_brain_correlation_cpgs.csv")
)
reliable_df <- inner_join(
  reliable_df, london.corr
)
reliable_df <- reliable_df %>% mutate(
  beta_spearman_cor_group = ifelse(
  beta_spearman_cor < 0.4, "Low", ifelse(beta_spearman_cor < 0.75 & beta_spearman_cor > 0.4, "Medium", "High")),
  residual_spearman_cor_group = ifelse(
  residual_spearman_cor < 0.4, "Low", ifelse(residual_spearman_cor < 0.75 & residual_spearman_cor > 0.4, "Medium", "High")))
```

```{r eval= F,echo = F}
write_csv(reliable_df, file.path(dir.reliable.score,
                                 "ADNI_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL_brain_blood_corr.csv"))
```

```{r eval= F,echo = F}
reliable_df <- read_csv( file.path(dir.reliable.score,
                                   "ADNI_ICC_beta_and_M_results_ind_diff_plates_samples_annotated_with_ADNI_AIBL_brain_blood_corr.csv"))
```


## Beta correlation vs ICC and HoLA 

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_df,
  y = "beta_spearman_cor",
  x = "ICC_Beta",
  xlab = "ICC",
  ylab = "London Beta correlation",
  size = .1,
  cor.coef = T,
  cor.method = "spearman"
)
```

```{r echo = F, fig.width=9, fig.height=8}
reliable_df$beta_spearman_cor_group <- factor(reliable_df$beta_spearman_cor_group, levels = c("Low", "Medium", "High"))
p_sum <- reliable_df %>% dplyr::group_by(beta_spearman_cor_group) %>% dplyr::summarise(n = dplyr::n())
ggboxplot(
  reliable_df,
  x = "beta_spearman_cor_group",
  y = "ICC_Beta",
  ylab = "ICC",
  xlab = "brain-blood correlation of DNA methylation levels",
  fill = "beta_spearman_cor_group",
  palette = "jco",
  #outlier.shape = NA
) + guides(fill = "none")
```

```{r eval = F}
ggsave(filename = file.path(dir.plot, "ADNI_ICC_with_brain_blood_corr_boxplot.pdf"),
       width = 8, height = 5)
```

```{r warning = F, fig.width=9,fig.height=5, echo = F}
ggboxplot(
  data = reliable_df, 
  x = "classification_Beta",
  y = "beta_spearman_cor",
  ylab = "London Beta correlation",
  xlab = "ICC group",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA
)
```

## Residuals correlation vs ICC

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_df,
  y = "residual_spearman_cor",
  x = "ICC_Beta",
  xlab = "ICC",
  ylab = "London residual correlation",
  size = .1,
  cor.coef = T,
  cor.method = "spearman"
)
```

```{r warning = F, fig.width=9,fig.height=5, echo = F}
ggboxplot(
  data = reliable_df, 
  x = "classification_Beta",
  y = "residual_spearman_cor",
  ylab = "London residual correlation",
  xlab = "ICC group",
  fill = "classification_Beta",
  palette = "jco",
  outlier.shape = NA
)
```

# Demenstrate that CpGs have higher probes reliability (mICC) are more likely located within DMRs 

```{r message = F}
DMR <- readxl::read_xlsx(file.path(dir.mike, "cpg_plus3.xlsx"))
```

## Combined sex analysis

```{r warning = F, fig.width=9,fig.height=5, echo = F}
ggboxplot(
  data = DMR %>% mutate(located_in_combined = ifelse(located_in_combined == 1, "Yes", "No")), 
  y = "ICC_Beta",
  x = "located_in_combined",
  ylab = "ICC",
  xlab = "probes located within AD-associated DMRs",
  fill = "located_in_combined",
  palette = "jco",
  outlier.shape = NA
) + guides(fill = "none")
ggsave(filename = file.path(dir.plot, "ADNI_ICC_with_combined_sex_DMR_analysis_boxplot.pdf"),
       width = 8, height = 5)
```

## Female

```{r warning = F, fig.width=9,fig.height=5, echo = F}
ggboxplot(
  data = DMR %>% mutate(located_in_females = ifelse(located_in_females == 1, "Yes", "No")), 
  y = "ICC_Beta",
  x = "located_in_females",
  ylab = "ICC",
  xlab = "probes located within AD-associated DMRs",
  fill = "located_in_females",
  palette = "jco",
  outlier.shape = NA
) + guides(fill = "none")
ggsave(filename = file.path(dir.plot, "Female_ADNI_ICC_with_combined_sex_DMR_analysis_boxplot.pdf"),
       width = 8, height = 5)
```

## Male

```{r warning = F, fig.width=9,fig.height=5, echo = F}
ggboxplot(
  data = DMR %>% mutate(located_in_males = ifelse(located_in_males == 1, "Yes", "No")), 
  y = "ICC_Beta",
  x = "located_in_males",
  ylab = "ICC",
  xlab = "probes located within AD-associated DMRs",
  fill = "located_in_males",
  palette = "jco",
  outlier.shape = NA
) + guides(fill = "none")
ggsave(filename = file.path(dir.plot, "Male_ADNI_ICC_with_combined_sex_DMR_analysis_boxplot.pdf"),
       width = 8, height = 5)
```

# Session Information

```{r}
devtools::session_info()
```
