---
title: "Critical evaluation of the reliability of DNA methylation probes on the Illumina MethylationEPIC BeadChip microarrays for dementia research"
subtitle: "ADNI reliable probes (ICC) property study"
author: "Wei Zhang, Lily Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: yeti
    toc: true
    number_sections: true
    df_print: paged
    code_download: false
    toc_float:
      collapsed: yes
    toc_depth: 3
    font_size: 10
editor_options:
  chunk_output_type: inline 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(gmodels)
library(SummarizedExperiment)

dir.base <- "../../"
dir.data.new <- file.path(dir.base, "data/ADNI")
dir.data.new.processed <- file.path(dir.data.new, "DNA_methylation/processed")
dir.data.supp <- file.path(dir.base, "data/SuppFiles")
dir.reliable.score <- file.path(dir.base, "analysis-results/reliable_score_different_plate")
dir.plot <- file.path(dir.base, "analysis-results/plots")

dir.extra <- file.path(dir.base, "external_files")
dir.extra.aux <- file.path(dir.extra,"Aux_Sync")
dir.lissette <- file.path(dir.base, "Lissette")
for(p in grep("dir",ls(),value = T)) dir.create(get(p),recursive = TRUE,showWarnings = FALSE)
```

```{r}
# Load auxilary function
source("../Utility.R")
```

# ICC compute

## Load results

### ICC results from ADNI replications

```{r message = F}
## Beta
reliable_score.beta <- read_csv(
  file.path(dir.reliable.score, "ADNI_sesame_icc_results_ind_measure_single_different_plate.csv")
) %>% dplyr::mutate(classification = factor(ifelse(value < 0.4, "Poor", 
                                            ifelse(value >= 0.4 & value < 0.6, "Fair",
                                                   ifelse(value >= 0.6 & value < 0.75, "Good", "Excellent"))),
                                            levels = c("Poor", "Fair", "Good", "Excellent"))) %>% 
  rename_with(.fn = function(x) paste0(x, "_Beta"), .cols = -contains(c("cpg")))
colnames(reliable_score.beta)[2] <- "ICC_Beta"
# M
reliable_score.M <- read_csv(
  file.path(dir.reliable.score, "ADNI_sesame_icc_results_ind_measure_single_M_values_different_plate.csv")
) %>% dplyr::mutate(classification = factor(ifelse(value < 0.4, "Poor", 
                                            ifelse(value >= 0.4 & value < 0.6, "Fair",
                                                   ifelse(value >= 0.6 & value < 0.75, "Good", "Excellent"))),
                                            levels = c("Poor", "Fair", "Good", "Excellent"))) %>%
  rename_with(.fn = function(x) paste0(x,"_M"), .cols = -contains(c("cpg")))
colnames(reliable_score.M)[2] <- "ICC_M"
```

### ICC results from Snugden et al.

```{r message = F}
## Load the pre-downloaded reliable CpGs suppliment file
reliable_CpG_Sugden <- readxl::read_xlsx(
  file.path(dir.data.supp, "1-s2.0-S2666389920300143-mmc2.xlsx"),
  skip = 2
) 
colnames(reliable_CpG_Sugden)[1] <- "cpg"
reliable_CpG_Sugden <- reliable_CpG_Sugden %>% dplyr::mutate(Sugden_classification = factor(ifelse(Reliability < 0.4, "Poor", 
                                            ifelse(Reliability >= 0.4 & Reliability < 0.6, "Fair",
                                                   ifelse(Reliability >= 0.6 & Reliability < 0.75, "Good", "Excellent"))),
                                            levels = c("Poor", "Fair", "Good", "Excellent"))) 
```

### Logue et al. (2017)

```{r message = F}
## Load the pre-downloaded reliable CpGs suppliment file
reliable_CpG_Logue <- readxl::read_xlsx(
  file.path(dir.data.supp, "Logue-ICC-values_Supplementary_Table_1.xlsx")
) 
colnames(reliable_CpG_Logue)[1] <- "cpg"
reliable_CpG_Logue <- reliable_CpG_Logue %>% dplyr::select(cpg, cor) %>% 
  dplyr::mutate(Logue_classification = factor(ifelse(cor < 0.4, "Poor", 
                                            ifelse(cor >= 0.4 & cor < 0.6, "Fair",
                                                   ifelse(cor >= 0.6 & cor < 0.75, "Good", "Excellent"))),
                                            levels = c("Poor", "Fair", "Good", "Excellent"))) 
```

### Bose et al. (2014)

```{r message = F}
## Load the pre-downloaded reliable CpGs suppliment file
reliable_CpG_Bose <- read_csv(
  file.path(dir.data.supp, "Bose-ARIC-ICC-values.csv")
) 
colnames(reliable_CpG_Bose)[1] <- "cpg"
reliable_CpG_Bose <- reliable_CpG_Bose %>% dplyr::select(cpg, `ICC value`) %>%
  dplyr::mutate(Bose_classification = factor(ifelse(`ICC value` < 0.4, "Poor", 
                                            ifelse(`ICC value` >= 0.4 & `ICC value` < 0.6, "Fair",
                                                   ifelse(`ICC value` >= 0.6 & `ICC value` < 0.75, "Good", "Excellent"))),
                                            levels = c("Poor", "Fair", "Good", "Excellent"))) 
```

## ADNI ICC of Beta and M values

```{r}
## Beta values
quantile(reliable_score.beta$ICC_Beta)
```

```{r warning = F, echo = F, fig.width=6,fig.height=4}
ggpubr::gghistogram(
  data = reliable_score.beta,
  x = "ICC_Beta",
  xlab = "ICC values",
  ylab = "count",
  combine = T,
  fill = "pink"
) + 
  geom_vline(aes(xintercept = mean(reliable_score.beta$ICC_Beta), 
                 colour = "mean ICC"), linetype = "dashed") +
  geom_vline(aes(xintercept = median(reliable_score.beta$ICC_Beta), 
                 colour = "median ICC"), linetype = "dashed") + 
  scale_colour_manual(breaks = c("mean ICC", "median ICC"),
                      values = c("mean ICC"="navy", "median ICC"="red"),
                      guide = guide_legend(title = ""))
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ADNI_ICC_distribution_different_plates_samples.pdf"),
  width = 8,
  height = 6
)
```

```{r}
## M values
quantile(reliable_score.M$ICC_M)
```

```{r warning = F, echo = F, fig.width=6,fig.height=4}
ggpubr::gghistogram(
  data = reliable_score.M,
  x = "ICC_M",
  xlab = "ICC M",
  ylab = "Frequency",
  combine = T,
  fill = "pink",
) + 
  geom_vline(aes(xintercept = 0.4, colour = "ICC = 0.4"), linetype = "dashed") +
  geom_vline(aes(xintercept = 0.6, colour = "ICC = 0.6"), linetype = "dashed") + 
  geom_vline(aes(xintercept = 0.75, colour = "ICC = 0.75"), linetype = "dashed") +
  scale_colour_manual(breaks = c("ICC = 0.4", "ICC = 0.6", "ICC = 0.75"),
                      values = c("ICC = 0.4"="red", "ICC = 0.6"="brown", 
                                 "ICC = 0.75"="black"),
                      guide = guide_legend(title = ""))
```

### Correlation

```{r}
reliable_score <- left_join(reliable_score.beta, reliable_score.M)
```

```{r echo = F, fig.width=9, fig.height=8}
ggscatter(
  reliable_score,
  x = "ICC_M",
  y = "ICC_Beta",
  xlab = "ICC M-values",
  ylab = "ICC Beta-values",
  size = .1,
  cor.coef = T,
  cor.method = "spearman",
)
```

### Agreement in classification

```{r}
Beta_classification <- reliable_score$classification_Beta
M_classification <- reliable_score$classification_M
```

```{r}
CrossTable(
  Beta_classification,
  M_classification,
  prop.c = FALSE, prop.r = FALSE, prop.t = TRUE, prop.chisq = FALSE, format = "SAS")
```

### Cohen's kappa of two ratings

```{r}
psych::cohen.kappa(cbind(Beta_classification, M_classification))
```

### Distribution of differences

```{r}
quantile(abs(reliable_score$ICC_Beta - reliable_score$ICC_M)) %>% round(digits = 4)
```

## ADNI ICC and other data ICC

### Correlation

```{r}
reliable_score.S <- full_join(reliable_CpG_Sugden, reliable_score.beta)
```

#### Snugden et al. ICC

```{r echo = F}
sum.df <- rbind(
  reliable_score.S %>% summarise(
  mean = mean(ICC_Beta, na.rm = T),
  median = median(ICC_Beta, na.rm = T),
  std = sd(ICC_Beta, na.rm = T)),
  reliable_score.S %>% summarise(
  mean = mean(Reliability, na.rm = T),
  median = median(Reliability, na.rm = T),
  std = sd(Reliability, na.rm = T))
) %>% mutate(study = c("ADNI", "Snugden"), .before = 1)
#write_csv(sum.df,file.path(dir.reliable.score, "Summary_table.csv"))
sum.df
```

```{r echo = F, fig.width=9, fig.height=8, warning=F}
ggscatter(
  reliable_score.S,
  x = "Reliability",
  y = "ICC_Beta",
  xlab = "ICC of Snugden et al",
  ylab = "ICC of ADNI",
  size = .1,
  cor.coef = T,
  cor.method = "spearman",
  cor.coef.coord = c(0.8, -0.3),
  cor.coeff.args = list(color = "navy", digits = 3)
)
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ICC_scatter_Snugden_vs_ADNI.pdf"),
  width = 9,
  height = 8
)
```

```{r}
reliable_score.l <- full_join(reliable_CpG_Logue, reliable_score.beta)
```

#### Logue et al. (2017) ICC

```{r echo = F, fig.width=9, fig.height=8, warning=F}
ggscatter(
  reliable_score.l,
  x = "cor",
  y = "ICC_Beta",
  xlab = "ICC of Logue et al.",
  ylab = "ICC of ADNI",
  size = .1,
  cor.coef = T,
  cor.method = "spearman",
  cor.coef.coord = c(0.7, -0.3),
  cor.coeff.args = list(color = "navy", digits = 3)
)
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ICC_scatter_Logue_vs_ADNI.pdf"),
  width = 9,
  height = 8
)
```

#### Bose et al. (2014) ICC

```{r}
reliable_score.B <- full_join(reliable_CpG_Bose, reliable_score.beta)
```

```{r echo = F, fig.width=9, fig.height=8, warning=F}
ggscatter(
  reliable_score.B,
  x = "ICC value",
  y = "ICC_Beta",
  xlab = "ICC of Bose et al.",
  ylab = "ICC of ADNI",
  size = .1,
  cor.coef = T,
  cor.method = "spearman",
  cor.coef.coord = c(0.7, -0.3),
  cor.coeff.args = list(color = "navy", digits = 3)
)
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ICC_scatter_Bose_vs_ADNI.pdf"),
  width = 9,
  height = 8
)
```

#### correlation table

```{r}
mat <- matrix(1, nrow = 3, ncol = 3)
colnames(mat) <- rownames(mat) <- c("Snugden", "Logue", "Bose")
#### cor of Snugden et al and Logue et al.
df <- inner_join(reliable_CpG_Sugden, reliable_CpG_Logue)
mat[1,2] <- mat[2,1] <- cor(df$Reliability, df$cor, method = "spearman")
df <- inner_join(reliable_CpG_Sugden, reliable_CpG_Bose)
mat[1,3] <- mat[3,1] <- cor(df$Reliability, df$`ICC value`, method = "spearman")
df <- inner_join(reliable_CpG_Logue, reliable_CpG_Bose)
mat[2,3] <- mat[3,2] <- cor(df$cor, df$`ICC value`, method = "spearman")
mat
#write_csv(mat %>% data.frame() %>% rownames_to_column("Study"), file.path(dir.reliable.score, "ICC_corr_other_study.csv"))
```



### Snugden et al. ICC

#### Agreement in classification

```{r}
ADNI_classification <- reliable_score.S$classification_Beta
Sugden_classification <- reliable_score.S$Sugden_classification
```

```{r}
CrossTable(
  ADNI_classification,
  Sugden_classification,
  prop.c = FALSE, prop.r = T, prop.t = F, prop.chisq = FALSE, format = "SAS")
Sugden_mat <- CrossTable(
  ADNI_classification,
  Sugden_classification,
  prop.c = FALSE, prop.r = T, prop.t = F, prop.chisq = FALSE, format = "SAS", dnn = c("ADNI", "Sugden"))
Sugden <- Sugden_mat$t
for(i in c(1:4)){
  for(j in c(1:4)){
    
    Sugden[i,j] <- paste0(Sugden_mat$t[i,j], " (", round(Sugden_mat$prop.row[i,j], 3) * 100, "%)")
    
  }
}
Sugden <- unclass(Sugden)
Sugden <- rbind(Sugden, colSums(Sugden_mat$t))
rownames(Sugden)[5] <- "Column Total"
Sugden <- data.frame(Sugden) %>% mutate("Row Total" =  c(rowSums(Sugden_mat$t),  Sugden[5,] %>% as.numeric() %>% sum))
```

#### Cohen's kappa of two ratings

```{r}
psych::cohen.kappa(cbind(ADNI_classification, Sugden_classification))
```

#### Distribution of differences

```{r}
quantile(abs(reliable_score.S$ICC_Beta - reliable_score.S$Reliability), na.rm = T) %>% round(digits = 4)
```

#### Wilcoxon test

```{r fig.width=6,fig.height=4, echo = F}
reliable_score.S %>% 
  dplyr::select(c("cpg", "ICC_Beta", "Reliability")) %>%
  pivot_longer(
    cols = c("ICC_Beta", "Reliability"),
    names_to = "Type",
    values_to = "ICC"
) %>% mutate(Type = ifelse(Type == "ICC_Beta", "ADNI", "Sugden")) %>% 
  ggboxplot(
    x = "Type",
    y = "ICC",
    fill = "Type",
    palette = "jco"
) + theme(axis.title.x = element_blank())#+ stat_compare_means(method = "wilcox.test", paired = F)

```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "Boxplot_Sugden_vs_ADNI.pdf"),
  width = 6,
  height = 4
)
```

```{r}
reliable_score.S <- reliable_score.S %>% 
  dplyr::select(c("cpg", "ICC_Beta", "Reliability")) %>%
  pivot_longer(
    cols = c("ICC_Beta", "Reliability"),
    names_to = "Type",
    values_to = "ICC"
) %>% mutate(Type = ifelse(Type == "ICC_Beta", "ADNI", "Sugden")) %>% na.omit()
summ.table(reliable_score.S, Group = "Type", type = "ICC")
```


### Bose et al. ICC

#### Agreement in classification

```{r}
ADNI_classification <- reliable_score.B$classification_Beta
Bose_classification <- reliable_score.B$Bose_classification
```

```{r}
CrossTable(
  ADNI_classification,
  Bose_classification,
  prop.c = FALSE, prop.r = T, prop.t = TRUE, prop.chisq = FALSE, format = "SPSS")
Bose_mat <- CrossTable(
  ADNI_classification,
  Bose_classification,
  prop.c = FALSE, prop.r = T, prop.t = F, prop.chisq = FALSE, format = "SAS", dnn = c("ADNI", "Bose"))
Bose <- Bose_mat$t
for(i in c(1:4)){
  for(j in c(1:4)){
    
    Bose[i,j] <- paste0(Bose_mat$t[i,j], " (", round(Bose_mat$prop.row[i,j], 3) * 100, "%)")
    
  }
}
Bose <- unclass(Bose)
Bose <- rbind(Bose, colSums(Bose_mat$t))
rownames(Bose)[5] <- "Column Total"
Bose <- data.frame(Bose) %>% mutate("Row Total" =  c(rowSums(Bose_mat$t),  Bose[5,] %>% as.numeric() %>% sum))
```

```{r eval = F}
writexl::write_xlsx(
  list(ADNI_Sugden = data.frame("Group" = rownames(Sugden), Sugden),
       ADNI_Bose = data.frame("Group" = rownames(Bose), Bose)),
  file.path(dir.reliable.score, "Cross_table_ADNI_ICC_vs_study.xlsx")
)
```


#### Cohen's kappa of two ratings

```{r}
psych::cohen.kappa(cbind(ADNI_classification, Bose_classification))
```

#### Distribution of differences

```{r}
quantile(abs(reliable_score.B$ICC_Beta - reliable_score.B$`ICC value`), na.rm = T) %>% round(digits = 4)
```

#### Wilcoxon test

```{r fig.width=6,fig.height=4, echo = F}
reliable_score.B %>% 
  dplyr::select(c("cpg", "ICC_Beta", "ICC value")) %>%
  pivot_longer(
    cols = c("ICC_Beta", "ICC value"),
    names_to = "Type",
    values_to = "ICC"
) %>% mutate(Type = ifelse(Type == "ICC_Beta", "ADNI", "Bose")) %>% 
  ggboxplot(
    x = "Type",
    y = "ICC",
    fill = "Type",
    palette = "jco"
) #+ stat_compare_means(method = "ks.test")
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "Boxplot_Bose_vs_ADNI.pdf"),
  width = 6,
  height = 4
)
```

```{r}
reliable_score.B <- reliable_score.B %>% 
  dplyr::select(c("cpg", "ICC_Beta", "ICC value")) %>%
  pivot_longer(
    cols = c("ICC_Beta", "ICC value"),
    names_to = "Type",
    values_to = "ICC"
) %>% mutate(Type = ifelse(Type == "ICC_Beta", "ADNI", "Bose")) %>% na.omit()
summ.table(reliable_score.B, Group = "Type", type = "ICC")
```

# Reproduce results in Snugden et al. (2020)

## Annotations

```{r eval = F}
### Auxillary function
load(file.path(dir.extra.aux,"great_EPIC_array_annotation.rda"))
load(file.path(dir.extra.aux,"E073_15_coreMarks_segments.rda"))
data <- readr::read_tsv(
  file.path(dir.extra,"AllPredictions.AvgHiC.ABC0.015.minus150.ForABCPaperV3.txt.gz")
)
CellType.selected <- readxl::read_xlsx(
  file.path(dir.extra, "Nassser study selected biosamples.xlsx"),col_names = FALSE
) %>% dplyr::pull(1)

annotate_results <- function(result){

  result$Islands.UCSC.Relation_to_Island <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19::Islands.UCSC[result$cpg,"Relation_to_Island"]
  result$UCSC_RefGene_Name <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19::Other[result$cpg,"UCSC_RefGene_Name"]       
  result$UCSC_RefGene_Group <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19::Other[result$cpg,"UCSC_RefGene_Group"]     
  result$UCSC_RefGene_Group[result$UCSC_RefGene_Group == ""] <- "Intergenic"
  
  result <- dplyr::left_join(result, great, by = "cpg")

  data.filtered <- data %>% dplyr::filter(CellType %in% CellType.selected) %>% 
    dplyr::filter(!isSelfPromoter)  %>% 
    dplyr::filter(class != "promoter")
  
  nasser.enhancer.gr <- data.filtered %>% makeGRangesFromDataFrame(
    start.field = "start",
    end.field = "end",
    seqnames.field = "chr",
    keep.extra.columns = TRUE
  )
  result.gr <- result %>% makeGRangesFromDataFrame(
      start.field = "start",
      end.field = "end",
      seqnames.field = "seqnames"
    )
  hits <- findOverlaps(result.gr, nasser.enhancer.gr) %>% as.data.frame()
  result$enhancer <- 0
  result$enhancer[unique(hits$queryHits)] <- 1
  
  return(result)
}
annotType <- IlluminaHumanMethylationEPICanno.ilm10b4.hg19::Manifest
```

```{r eval = F}
reliable_score.sum.annotated <- annotate_results(reliable_score)
annotTypeCompleteCol1 <- annotType[reliable_score.sum.annotated$cpg, ]
reliable_score.sum.annotated$Probe_Type <- annotTypeCompleteCol1$Type
```

## Relationship between reliability vs. mean and standard deviations in beta ICC

```{r eval = F}
### Load beta values
beta <- readRDS( file.path(dir.data.new.processed,"ADNI_RnBeads_and_pooBAH_filtered_beta_different_plates.RDS"))
beta.mean <- data.frame(beta.mean = rowMeans(beta, na.rm = T)) %>% rownames_to_column("cpg")
beta.sd <- data.frame(beta.sd = genefilter::rowSds(beta, na.rm = T)) %>% dplyr::mutate(cpg = rownames(beta), .before = 1)
reliable_score.sum.annotated <- left_join(
  reliable_score.sum.annotated, beta.mean, by = "cpg"
) %>% left_join(beta.sd, by = "cpg")
```

```{r eval = F}
set.seed(12345)
sd.rand <- apply(samples.id, 1, function(i) i[sample(2:3,1)])
sd.random <- apply(beta[,sd.rand], 1,sd)
sd.random <- data.frame(cpg = names(sd.random), beta.sd.rand = sd.random)
reliable_score.sum.annotated <- left_join(reliable_score.sum.annotated, sd.random)
```

```{r eval = F}
write_csv(
  reliable_score.sum.annotated,
    file.path(dir.reliable.score, "ADNI_icc_beta_and_M_results_ind_samples_annotated_different_plates.csv")
)
```

```{r echo = F, message=F}
reliable_score.sum.annotated <- read_csv(
  file.path(dir.reliable.score, "ADNI_icc_beta_and_M_results_ind_samples_annotated_different_plates.csv")
)
```

### Reliability vs mean

```{r echo = F}
df_stat <- list()
```

```{r fig.width=9,fig.height=5, echo = F}
p <- reliable_score.sum.annotated %>% 
  mutate(group.mean = ifelse(beta.mean < 0.2, "0 - 0.2",
                              ifelse(beta.mean < 0.4 & beta.mean >= 0.2, "0.2 - 0.4",
                                     ifelse(beta.mean < 0.6 & beta.mean >= 0.4, "0.4 - 0.6", 
                                            ifelse(beta.mean < 0.8 & beta.mean >= 0.6, "0.6 - 0.8", "0.8 - 1")))),
         group.mean = factor(group.mean, 
                              levels = c("0 - 0.2", 
                                         "0.2 - 0.4",
                                         "0.4 - 0.6",
                                         "0.6 - 0.8", 
                                         "0.8 - 1")))
p_sum <- p %>% dplyr::group_by(group.mean) %>% dplyr::summarise(n = dplyr::n())
ggboxplot(
  p,
  x = "group.mean",
  y = "ICC_Beta",
  ylab = "ICC",
  xlab = "Average beta value of the probe",
  fill = "group.mean",
  palette = "jco",
  outlier.shape = NA) + guides(fill="none")
# ) + annotate( "text", label = paste0("n_probes: ", p_sum$n[1]), x = 1, y = 1.2, size = 4) +
#   annotate( "text", label = paste0("n_probes: ", p_sum$n[2]), x = 2, y = 1.2, size = 4) +
#   annotate( "text", label = paste0("n_probes: ", p_sum$n[3]), x = 3, y = 1.2, size = 4) + 
#   annotate( "text", label = paste0("n_probes: ", p_sum$n[4]), x = 4, y = 1.2, size = 4) +
#   annotate( "text", label = paste0("n_probes: ", p_sum$n[5]), x = 5, y = 1.2, size = 4)
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ADNI_Boxplot_ICC_vs_Mean.pdf"),
  width = 8,
  height = 4
)
```

```{r echo = F}
df_stat[[1]] <- p %>% group_by(group.mean) %>% 
  dplyr::summarise(
    num_of_probes = n(),
    Mean = mean(ICC_Beta),
    SD = sd(ICC_Beta),
    Min = min(ICC_Beta),
    Median = median(ICC_Beta),
    Max = max(ICC_Beta)
  ) %>% print 
```

### Reliability vs sd

```{r fig.width=9,fig.height=5, echo = F}
q0 <- quantile(reliable_score.sum.annotated$beta.sd.rand)
q <- paste0("Q" , 1:4, ": ", 
            sapply(1:4, function(k) paste0(round(q0[k],2), "-", round(q0[k + 1],2))))
my_comparisons <- list( c(q[1], q[2]), c(q[2], q[3]), c(q[3], q[4]) )
df <- reliable_score.sum.annotated %>% 
  dplyr::mutate(Quantile.sd = ifelse(beta.sd.rand < q0[2], q[1],
                              ifelse(beta.sd.rand < q0[3] & beta.sd.rand >= q0[2], q[2],
                                     ifelse(beta.sd.rand < q0[4] & beta.sd.rand >= q0[3], q[3], q[4]))),
         Quantile.sd = factor(Quantile.sd, levels = q))
ggboxplot(
  df,
  x = "Quantile.sd",
  y = "ICC_Beta",
  ylab = "ICC",
  xlab = "Quartile of probe standard deviations",
  fill = "Quantile.sd",
  palette = "jco",
  outlier.shape = NA
) + guides(fill="none")
```

```{r eval = F, echo = F}
ggsave(
  filename = file.path(dir.plot, "ADNI_Boxplot_ICC_vs_SD.pdf"),
  width = 8,
  height = 4
)
```

```{r echo = F}
df_stat[[2]] <- df %>% group_by(Quantile.sd) %>% 
  dplyr::summarise(
    num_of_probes = n(),
    Mean = mean(ICC_Beta),
    SD = sd(ICC_Beta),
    Min = min(ICC_Beta),
    Median = median(ICC_Beta),
    Max = max(ICC_Beta)
  ) %>% print 
```

```{r eval = F,echo = F}
names(df_stat) <- c("ICC_vs_Mean","ICC_vs_SD")
writexl::write_xlsx(
  df_stat,
  file.path(dir.reliable.score, "ADNI_ICC_vs_beta_SD_Mean_table.xlsx")
)
```



## Relationship betwene ICCs vs Islands.UCSC.Relation_to_Island

First, we combine the shore and shelf

```{r}
df.list <- list()
reliable_score.sum.annotated <- reliable_score.sum.annotated %>%
  mutate(Relation_to_Island = ifelse(Islands.UCSC.Relation_to_Island %in% c("N_Shore", "S_Shore"), "Shore", 
                                     ifelse(Islands.UCSC.Relation_to_Island %in% c("N_Shelf", "S_Shelf"),
                                                                                   "Shelf",Islands.UCSC.Relation_to_Island)))
```

```{r fig.width=9,fig.height=5, echo = F}
reliable_score.sum.annotated %>% filter(!is.na(Relation_to_Island)) %>% 
  ggboxplot(
    x = "Relation_to_Island",
    y = "ICC_Beta",
    ylab = "ICC",
    xlab = "UCSC Relation to Island",
    fill = "Relation_to_Island",
    palette = "jco"
  ) + guides(fill = "none")
# ggsave(filename = file.path(dir.plot, "ADNI_probe_reliability_with_relation_to_island_boxplot.pdf"),
#        width = 8, height = 4)
```

```{r echo = F}
df.list[[1]] <- summ.table(df = reliable_score.sum.annotated, Group = "Relation_to_Island") %>% print
```

## ICCs vs. UCSC_RefGene_Group

```{r warning=F}
reliable_score.sum.annotated <- reliable_score.sum.annotated %>%
  separate(UCSC_RefGene_Group,into = "RefGene_Group", sep = ";", remove = F) %>%
  mutate(RefGene_Group = factor(RefGene_Group, levels = c("TSS1500", "TSS200", "5'UTR", "1stExon", "ExonBnd", "Body", "3'UTR", "Intergenic")))
```

```{r fig.width=9,fig.height=5, echo = F}
reliable_score.sum.annotated %>% filter(!is.na(RefGene_Group)&RefGene_Group != "") %>% 
  ggboxplot(
    x = "RefGene_Group",
    y = "ICC_Beta",
    ylab = "ICC",
    xlab = "UCSC RefGene Group",
    fill = "RefGene_Group",
    palette = "jco"
  ) + guides(fill = "none")
ggsave(filename = file.path(dir.plot, "ADNI_probe_reliability_with_RefGene_Group_boxplot.pdf"),
       width = 12, height = 5)
```

```{r echo = F}
df.list[[2]] <- summ.table(df = reliable_score.sum.annotated, Group = "RefGene_Group") %>% print
```

## ICCs vs. type I or type II probes

```{r fig.width=6,fig.height=5, echo = F}
reliable_score.sum.annotated %>% filter(!is.na(Probe_Type)) %>% 
  ggboxplot(
    x = "Probe_Type",
    y = "ICC_Beta",
    ylab = "ICC",
    xlab = "Probe Type",
    fill = "Probe_Type",
    palette = "jco"
  ) + guides(fill = "none")
# ggsave(filename = file.path(dir.plot, "ADNI_probe_reliability_with_probe_type_boxplot.pdf"),
#        width = 8, height = 5)
```

```{r echo = F}
df.list[[3]] <- summ.table(df = reliable_score.sum.annotated, Group = "Probe_Type") %>% print
```

## ICCs vs. Enhancer

```{r fig.width=6,fig.height=5, echo = F}
reliable_score.sum.annotated %>% filter(!is.na(Probe_Type)) %>% 
  ggboxplot(
    x = "enhancer",
    y = "ICC_Beta",
    ylab = "Probe Reliabiility",
    xlab = "Enhancer",
    fill = "enhancer",
    palette = "jco"
  ) + stat_compare_means(method = "wilcox.test")
# ggsave(filename = file.path(dir.plot, "ADNI_probe_reliability_with_Enhancer_boxplot.pdf"),
#        width = 8, height = 6)
```

```{r echo = F}
df.list[[4]] <- summ.table(df = reliable_score.sum.annotated, Group = "enhancer") %>% print
```

## ICC vs mQTLs

```{r message = F}
mQTLs <- readxl::read_xlsx(
  file.path(dir.lissette, "cpg_mqtl_reliabilityAnalysis_06022023.xlsx")
)
reliable_df <- left_join(reliable_score.sum.annotated, mQTLs)
```

```{r echo = F}
reliable_df  %>% filter(!is.na(mqtl)) %>%
  ggboxplot(
    x = "mqtl",
    y = "ICC_Beta",
    ylab = "Reliabiility",
    xlab = "mQTLs",
    fill = "mqtl",
    palette = "jco"
  ) + stat_compare_means(method = "wilcox.test")
# ggsave(filename = file.path(dir.plot, "ADNI_probe_reliability_with_mQTLs_boxplot.pdf"),
#        width = 8, height = 6)
```

```{r echo = F}
df.list[[5]] <- summ.table(df = reliable_df, Group = "mqtl") %>% print
```

```{r echo = F, eval = F}
names(df.list) <- c(
  "ICC_vs_Relation_to_Island",
  "ICC_vs_RefGene_Group",
  "ICC_vs_Probe_Type",
  "ICC_vs_Enhancer",
  "ICC_vs_mQTLs"
)
writexl::write_xlsx(
  df.list,
  file.path(dir.reliable.score, "ADNI_ICC_vs_CpG_groups.xlsx")
)
```

## Pie Chart

```{r echo = F, fig.width=5, fig.height=6}
reliable_score.sum.annotated %>% filter(Probe_Type == "I") %>%
  group_by(Relation_to_Island) %>%
  summarise(N = n()) %>% 
  ggpie(
    x = "N",
    group = "Relation_to_Island",
    fill = "Relation_to_Island",
    palette = "jco",
    color = "white",
    main = "Probe Type I"
  )
reliable_score.sum.annotated %>% filter(Probe_Type == "I") %>%
  group_by(RefGene_Group) %>%
  summarise(N = n()) %>% 
  ggpie(
    x = "N",
    group = "RefGene_Group",
    fill = "RefGene_Group",
    palette = "jco",
    color = "white",
    main = "Probe Type I"
  )
```

```{r echo = F, fig.width=5, fig.height=6}
reliable_score.sum.annotated %>% filter(Probe_Type == "II") %>%
  group_by(Relation_to_Island) %>%
  summarise(N = n()) %>% 
  ggpie(
    x = "N",
    group = "Relation_to_Island",
    fill = "Relation_to_Island",
    palette = "jco",
    color = "white",
    main = "Probe Type II"
  )
reliable_score.sum.annotated %>% filter(Probe_Type == "II") %>%
  group_by(RefGene_Group) %>%
  summarise(N = n()) %>% 
  ggpie(
    x = "N",
    group = "RefGene_Group",
    fill = "RefGene_Group",
    palette = "jco",
    color = "white",
    main = "Probe Type II"
  )
```





# Session Information

```{r}
devtools::session_info()
```
