---
title: "Critical evaluation of the reliability of DNA methylation probes on the Illumina MethylationEPIC BeadChip microarrays for dementia research"
subtitle: "ADNI SeSame 2 normalization"
author: "Wei Zhang, Lily Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: yeti
    toc: true
    number_sections: true
    df_print: paged
    code_download: false
    toc_float:
      collapsed: yes
    toc_depth: 3
    font_size: 10
editor_options:
  chunk_output_type: inline 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
library(tidyverse)
library(SummarizedExperiment)
library(ExperimentHub)
library(irr)
library(sesame)
library(RnBeads)
#sesameDataCache() # Please cache SeSAMe data if you're first time using sesame package.
dir.base <- "../../"
dir.data.new <- file.path(dir.base, "data/ADNI/")
dir.raw.selected <-  file.path(dir.data.new, "/DNA_methylation/raw/")
dir.RnBeads.selected <- file.path(dir.data.new, "/DNA_methylation/RnBeads/")
dir.data.new.processed <- file.path(dir.data.new, "/DNA_methylation/processed/")
dir.reliable.score <- file.path(dir.base, "/analysis-results/reliable_score_different_plate/")
for(p in grep("dir",ls(),value = T)) dir.create(get(p),recursive = TRUE,showWarnings = FALSE)
```



# Load data

```{r message = F}
samples.duplicated.all <- readxl::read_xlsx(
  file.path(dir.data.new.processed, "ADNI_DNA_Methylation_Sample_Duplicated_IND.xlsx")
)
## Select samples with different plates
samples.diff.plate.slide <- samples.duplicated.all %>% 
  dplyr::select(RID, PlateNumber) %>% unique %>% group_by(RID) %>% 
  dplyr::summarise(n = n()) %>% filter(n == 2)
nrow(samples.diff.plate.slide) # 69
samples.duplicated.ind <- samples.duplicated.all %>% filter(RID %in% samples.diff.plate.slide$RID)
samples.duplicated.same.plate <- samples.duplicated.all %>% filter(!RID %in% samples.diff.plate.slide$RID)

writexl::write_xlsx(samples.duplicated.ind, file.path(dir.data.new.processed, "ADNI_DNA_Methylation_Sample_Duplicated_Diff_Plates.xlsx"))
writexl::write_xlsx(samples.duplicated.same.plate, file.path(dir.data.new.processed, "ADNI_DNA_Methylation_Sample_Duplicated_Same_Plates.xlsx"))
```

```{r}
rnbsSaveDir <- file.path(dir.data.new.processed, "ADNI_Rnb_file_on_ind_dup_samples")
rnb.set <- load.rnb.set(rnbsSaveDir)
samples.duplicated.same.plate <- which(rnb.set@pheno$Sample_ID %in% samples.duplicated.same.plate$barcodes)
rnb.set.filtered <- RnBeads::remove.samples(rnb.set, samples.duplicated.same.plate)
rnbsSaveDir <- file.path(dir.data.new.processed, "ADNI_Rnb_file_on_ind_dup_samples_no_samee_plate")
save.rnb.set(rnb.set.filtered, rnbsSaveDir, archive=FALSE)
```

```{r}
rm(rnb.set)
```



# Preprocessing with Sesame 2 pipeline

```{r eval = F}
rnbsSaveDir <- file.path(dir.data.new.processed, "ADNI_Rnb_file_on_ind_dup_samples_no_samee_plate")
rnb.set.filtered <- load.rnb.set(rnbsSaveDir)
```


## QC step

1. remove probes if their target sequences overlap with a SNP at any base
2. remove probes outside of CpG context
3. remove known cross-reactive probe
4. remove X/Y probe
5. use the iterative Greedy cut algorithm to filter out samples and probes, using a detection p value threshold of 0.01
6. remove probes if more than 5% of the samples had a missing value

### 1. removed probes if their target sequences overlap with a SNP at any base

```{r eval = F}
filterRes <- rnb.execute.snp.removal(rnb.set.filtered, snp = "any")
rnbs.filtered.snq <- filterRes$dataset
probes.raw <- nsites(rnb.set.filtered) # 866895
probes.after.remove.snp <- nsites(rnbs.filtered.snq) # 727174
```

### 2. remove probes outside of CpG context

```{r eval = F}
filterRes <- rnb.execute.context.removal(rnbs.filtered.snq)
rnbs.filtered.cpg <- filterRes$dataset
probes.after.remove.non.cpgs <- nsites(rnbs.filtered.cpg) # 724560
```

### 3. remove known cross-reactive probe

```{r eval = F}
filterRes <- rnb.execute.cross.reactive.removal(rnbs.filtered.cpg)
rnbs.filtered.cross.reactive <- filterRes$dataset
probes.after.remove.cross.reactive <- nsites(rnbs.filtered.cross.reactive) # 691714
```

### 4. remove X/Y probe

```{r eval = F}
filterRes <- rnb.execute.sex.removal(rnbs.filtered.cross.reactive)
rnbs.filtered.sex <- filterRes$dataset
probes.after.remove.sex <- nsites(rnbs.filtered.sex) # 674810
```

### 5. use the iterative Greedy cut algorithm to filter out samples and probes, using a detection p value threshold of 0.01

```{r eval = F}
filterRes <- rnb.execute.greedycut(
  rnbs.filtered.sex,
  pval.threshold = 0.01
)
rnbs.filtered.p <- remove.samples(remove.sites(rnbs.filtered.sex, filterRes$sites), filterRes$samples)
probes.after.remove.p <- nsites(rnbs.filtered.p) # 661308
```

### 6. remove probes if more than 5% of the samples had a missing value

```{r eval = F}
filterRes <- rnb.execute.na.removal(rnbs.filtered.p, threshold = 0.05)
rnbs.filtered.na <- filterRes$dataset
probes.after.remove.na <- nsites(rnbs.filtered.na) # 661304
```

```{r eval = F}
rnbsSaveDir <- file.path(dir.data.new.processed, "ADNI_Rnb_file_on_ind_dup_samples_filtered_QC_different_plate")
save.rnb.set(rnbs.filtered.na, rnbsSaveDir, archive=FALSE)
df.probes.remove.RnBeads <- data.frame(
  "Description" = c(
    "Raw", "Removing overlap with a SNP at any base",
    "Removing probes outside of CpG context",
    "Removing any known cross-reactive probe",
    "Removing X/Y probe",
    "Removing probes using Greedy cut algorithm at 0.01 pvalue threshold",
    "Removing probes if more than 5% of the samples had a missing value"
  ),
  "Number of probes" = c(
    probes.raw,
    probes.after.remove.snp,
    probes.after.remove.non.cpgs,
    probes.after.remove.cross.reactive,
    probes.after.remove.sex,
    probes.after.remove.p,
    probes.after.remove.na
  )
)
```

```{r eval = F}
beta.RnBeads <- meth(rnbs.filtered.na , row.names = T)
saveRDS(
  beta.RnBeads,
  file.path(dir.data.new.processed, "ADNI_RnBeads_filtered_beta_different_plates.RDS")
)
```

```{r}
beta.RnBeads <- readRDS(file.path(dir.data.new.processed, "ADNI_RnBeads_filtered_beta_different_plates.RDS"))
probes.masked.by.RnBeads <- (sites(rnb.set.filtered) %>% rownames())[-which((sites(rnb.set.filtered) %>% rownames()) %in% rownames(beta.RnBeads))] #205591
```



```{r}
rm(rnb.set.filtered, filterRes, rnbs.filtered.cpg, rnbs.filtered.cross.reactive, rnbs.filtered.p, rnbs.filtered.sex, rnbs.filtered.snq)
gc()
```


### Convert to sigDF

```{r}
samples.duplicated.ind <- readxl::read_xlsx( file.path(dir.data.new.processed, "ADNI_DNA_Methylation_Sample_Duplicated_Diff_Plates.xlsx"))
```


```{r eval = F}
idat.file <- list.files(
  path = dir.raw.selected,
  pattern = paste0(samples.duplicated.ind$barcodes, collapse = "|"),
  full.names = T
)  %>%
  gsub("_Grn.idat|_Red.idat", "", .) %>%
  unique
doParallel::registerDoParallel(10)
sdfs.list <- plyr::llply(
  idat.file,
  .fun = function(idat){
    readIDATpair(
      prefix.path = idat,
      platform = "EPIC"
    )
  },.parallel = F
)
names(sdfs.list) <- list.files(
  path = dir.raw.selected,
  pattern = paste0(samples.duplicated.ind$barcodes, collapse = "|"),
  full.names = F
)  %>%
  gsub("_Grn.idat|_Red.idat", "", .) %>%
  unique
nprobe.raw <- sdfs.list$`200223270011_R01C01` %>% nrow() # 866553
```

### pOOBAH masking


```{r eval = F}
doParallel::registerDoParallel(10)
sdfs.pooBAH <- plyr::llply(
  sdfs.list,
  .fun = function(sdfs){
    pOOBAH(
      sdf = sdfs
    )
  }, .parallel = T
)
```

```{r eval = F}
## Add mask for 20% samples missing
probes.pooBAH.mask <- plyr::llply(
  sdfs.pooBAH,
  .fun = function(sdfs){
    sdfs$Probe_ID[sdfs$mask == T]
  }
) %>% unlist %>% unique()
length(probes.pooBAH.mask) # 224216
beta.pooBAH.mask <- plyr::llply(
  sdfs.pooBAH,
  .fun = function(sdfs){
    getBetas(sdfs)
  }
) %>% Reduce(cbind,.)
colnames(beta.pooBAH.mask) <- names(sdfs.pooBAH)
probes.masked.by.poobah <- rownames(beta.pooBAH.mask)[which(rowSums(is.na(beta.pooBAH.mask))/ncol(beta.pooBAH.mask) >= 0.2)] #42263
df.probes.remove.RnBeads <- rbind(
  df.probes.remove.RnBeads,
  c("Removing additional probes if more than 20% of samples masked by pOOBAH",df.probes.remove.RnBeads$Number.of.probes[7] -  length(probes.masked.by.poobah[which(!probes.masked.by.poobah %in% probes.masked.by.RnBeads)]))
) #640985
```

### Add mask from QC RnBeads and pOOBAH

```{r eval = F}
doParallel::registerDoParallel(10)
sdfs.from.qc <- plyr::llply(
  sdfs.list,
  .fun = function(sdfs){
    sdfs <- prefixMaskButCG(sdfs)
    addMask(
      sdf = sdfs,
      probes = union(probes.masked.by.RnBeads,probes.masked.by.poobah)
    )
  }, .parallel = T
)
```

### Remove masked probes

```{r eval = F}
doParallel::registerDoParallel(10)
sdfs.no.mask <- plyr::llply(
  sdfs.from.qc,
  .fun = function(sdfs){
    noMasked(
      sdf = sdfs
    )
  }, .parallel = T
)

save(sdfs.no.mask, file = file.path(dir.data.new.processed,"ADNI_RnBeads_and_pooBAH_filtered_sdf_different_plates.rda"))
```

```{r eval = F}
rm(sdfs.from.qc, sdfs.pooBAH, sdfs.list)
gc()
```


## Normalization

```{r eval = F}
load(file.path(dir.data.new.processed,"ADNI_RnBeads_and_pooBAH_filtered_sdf_different_plates.rda"))
```


```{r eval = F}
### Noob
doParallel::registerDoParallel(10)
sdfs.noob <- plyr::llply(
  sdfs.no.mask,
  .fun = function(sdfs){
    noob(
      sdf = sdfs
    )
  }, .parallel = T
)
### nonlinear dye bias correction
sdfs.dyeBiasNL <- plyr::llply(
  sdfs.noob,
  .fun = function(sdfs){
    dyeBiasNL(
      sdf = sdfs
    )
  }, .parallel = T
)

save(sdfs.dyeBiasNL, file = file.path(dir.data.new.processed,"ADNI_RnBeads_and_pooBAH_filtered_sdf_normalized_different_plates.rda"))
```


```{r eval = F}
## Get beta
doParallel::registerDoParallel(10)
beta.final <- plyr::llply(
  sdfs.dyeBiasNL,
  .fun = function(sdfs){
    getBetas(sdfs)
  },.parallel = T
) %>% Reduce(cbind,.)

colnames(beta.final) <- names(sdfs.dyeBiasNL)

df.probes.remove.RnBeads <- rbind(
  df.probes.remove.RnBeads,
  c("Removing additional non-cg probes in SigDF", nrow(beta.final))
)

saveRDS(
  beta.final,
  file.path(dir.data.new.processed, "ADNI_RnBeads_and_pooBAH_filtered_beta_different_plates.RDS")
)

write_csv(
  df.probes.remove.RnBeads,
  file.path(dir.data.new.processed, "ADNI_different_plates_probes_removed_in_each_step.csv")
)
```

```{r}
rm(beta.final)
gc()
```


# Calculate reliability


```{r eval = F}
beta <- readRDS( file.path(dir.data.new.processed,"ADNI_RnBeads_and_pooBAH_filtered_beta_different_plates.RDS"))
samples.duplicated.all <- readxl::read_xlsx(
  file.path(dir.data.new.processed, "ADNI_DNA_Methylation_Sample_Duplicated_IND.xlsx")
)
samples.duplicated.ind <- samples.duplicated.all %>% filter(RID %in% samples.diff.plate.slide$RID)
samples.id <- samples.duplicated.ind %>% dplyr::select(RID_Phase_Edata, barcodes) %>% 
  dplyr::group_by(RID_Phase_Edata) %>% 
  dplyr::mutate(obs = dplyr::row_number()) %>% 
  pivot_wider(names_from = "obs", 
              values_from = "barcodes", 
              names_prefix="barcodes")
```

```{r eval = F}
## Auxillary function
create_rating <- function(samples.id, beta, model = "twoway", type = "agreement", unit = "average", random = F, scale = F){
  
  barcodes <- samples.id[,c("barcodes1", "barcodes2")]
  if(random){
    sam <- sample.int(size = nrow(barcodes), n = 2, replace = T)
    barcodes <- sapply(1:nrow(barcodes),  function(i) c(barcodes[i,sam[i]],barcodes[i,-sam[i]])) %>% t() 
    rownames(barcodes) <- samples.id$RID_Phase_Edata
  }

  
  doParallel::registerDoParallel(20)
  rating_matrix <- plyr::adply(
    beta,
    .margins = 1,
    .fun = function(probe){
      rating <- plyr::aaply(
        barcodes,
        .margins = 1, 
        .fun = function(bar){
          probe[bar %>% unlist()]
        },.expand = F
      )
      if(scale) rating <- scale(rating)
      ic <- icc(rating, model = model, type = type, unit = unit)
      data.frame(value = ic$value, r0 = ic$r0, Fvalue = ic$Fvalue,
                 df1 = ic$df1, df2 = ic$df2, p.value = ic$p.value, 
                 lbound = ic$lbound, ubound = ic$ubound)
    },.parallel = T, .id = "cpg"
  )
  
  return(rating_matrix)
}
```

```{r eval = F}
## Beta
reliable_score_single <- create_rating(samples.id, beta, unit = "single")
colnames(reliable_score_single)[1] <- "cpg"
## M
M <- minfi::logit2(beta)
reliable_score_single.M <- create_rating(samples.id, M, unit = "single")
colnames(reliable_score_single.M)[1] <- "cpg"
```


```{r eval = F}
write_csv(
  reliable_score_single %>% arrange(desc(value)),
  file.path(dir.reliable.score, "ADNI_sesame_icc_results_ind_measure_single_different_plate.csv")
)
write_csv(
  reliable_score_single.M %>% arrange(desc(value)),
  file.path(dir.reliable.score, "ADNI_sesame_icc_results_ind_measure_single_M_values_different_plate.csv")
)
```


# Session Information

```{r}
devtools::session_info()
```

